---
title: "iNat 2021 - Graz City Challenge"
author: "Hannes Oberreiter"
date: "5/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

library(tidyverse)
library(here)
library(kableExtra)

data <- read_csv("data/observations-149633.csv")
```

```{r}
distinctSpeciesNumber <- data %>% 
  drop_na(taxon_genus_name) %>% 
  distinct(scientific_name) %>% 
  nrow()

statsObserver <- data %>% 
  group_by(user_id) %>%
  summarise(
    n = n(),
    mean = mean(n),
    median = median(n)
  )

singleSpeciesFound <- data %>% 
  drop_na(taxon_species_name) %>% 
  count(taxon_species_name) %>% 
  filter(n == 1)
```

## City Nature Challenge 2021: Graz

Date: 30. April 2021 - 3. May 2021

- Observations: `r nrow(data)`
- Distinct Species: `r distinctSpeciesNumber``
- Distinct Observer: `r nrow(statsObserver)`
- Observer - Mean / Median Observations: `r round(mean(statsObserver$mean),1)` / `r median(statsObserver$median)`
- Total number of Unique Observations (Species Level): `r nrow(singleSpeciesFound)`

### Determination Level

```{r}
data <- data %>% mutate(
  det = case_when(
    !is.na(taxon_species_name) ~ "Species",
    !is.na(taxon_genus_name) ~ "Genus",
    !is.na(taxon_family_name) ~ "Family",
    !is.na(taxon_order_name) ~ "Order",
    TRUE ~ "Phylum/Unknown"
  )
)
data %>% group_by(det) %>% 
  summarize(
    n = n(),
    np = n()/nrow(data) * 100,
    .groups = 'drop'
  ) %>% 
  arrange(desc(np)) %>% 
  mutate(name=factor(det, levels=det)) %>% 
  ggplot() +
  aes(x = name, y = np, fill = name) +
  geom_bar(position = 'dodge', stat='identity', show.legend = FALSE) +
  geom_text(aes(label=paste("n =", n)), position=position_dodge(width=0.9), hjust=-0.25) +
  xlab("") + ylab("Observations [%]") + ylim(0, 100) +
  coord_flip() +
  theme_classic()
```

### Icon Taxa Classes

```{r}
data %>% group_by(iconic_taxon_name) %>% 
  summarize(
    n = n(),
    np = n()/nrow(data) * 100,
    .groups = 'drop'
  ) %>% 
  arrange(desc(np)) %>% 
  mutate(name=factor(iconic_taxon_name, levels=iconic_taxon_name))  %>%   # This trick update 
  ggplot() +
  aes(x = name, y = np, fill = name) +
  geom_bar(position = 'dodge', stat='identity', show.legend = FALSE) +
  geom_text(aes(label=paste("n =", n)), position=position_dodge(width=0.9), hjust=-0.25) +
  xlab("") + ylab("Observations [%]") + ylim(0, 100) +
  coord_flip() +
  theme_classic()
```

#### Plantae Order

```{r, fig.height=10}
data %>% filter(iconic_taxon_name == "Plantae") %>%
  drop_na(taxon_order_name) %>% 
  group_by(taxon_order_name) %>%
  summarize(
    n = n(),
    np = n()/nrow(data) * 100,
    .groups = 'drop'
  ) %>% 
  arrange(desc(np)) %>% 
  mutate(name=factor(taxon_order_name, levels=taxon_order_name))  %>%   # This trick update 
  ggplot() +
  aes(x = name, y = np, fill = name) +
  geom_bar(position = 'dodge', stat='identity', show.legend = FALSE) +
  geom_text(aes(label=paste("n =", n)), position=position_dodge(width=0.9), hjust=-0.25) +
  xlab("") + ylab("Observations [%]") + ylim(0, 100) +
  coord_flip() +
  theme_classic()
```

#### Insecta Order

```{r, fig.height=5}
data %>% filter(iconic_taxon_name == "Insecta") %>%
  drop_na(taxon_order_name) %>% 
  group_by(taxon_order_name) %>%
  summarize(
    n = n(),
    np = n()/nrow(data) * 100,
    .groups = 'drop'
  ) %>% 
  arrange(desc(np)) %>% 
  mutate(name=factor(taxon_order_name, levels=taxon_order_name))  %>%   # This trick update 
  ggplot() +
  aes(x = name, y = np, fill = name) +
  geom_bar(position = 'dodge', stat='identity', show.legend = FALSE) +
  geom_text(aes(label=paste("n =", n)), position=position_dodge(width=0.9), hjust=-0.25) +
  xlab("") + ylab("Observations [%]") + ylim(0, 100) +
  coord_flip() +
  theme_classic()
```



### Birds - Annex I

```{r}
birds <- read_csv2("data/birds.csv") %>% 
  rename_with(~"scientific_name", 1) %>% 
  select(scientific_name, `Annex I`) %>% 
  rename(AnnexI = `Annex I`) %>% 
  drop_na(AnnexI) %>% 
  right_join(data, by = c("scientific_name")) %>% 
  drop_na(AnnexI) %>% 
  count(scientific_name, common_name, AnnexI) %>% 
  arrange(scientific_name)

kbl(birds, booktabs = TRUE) %>% 
  kable_classic("hover", full_width = F, html_font = "Helvetica Neue")
```

### Other - Annex II und IV

```{r}
# Load Species List to extract Anhang II und IV Arten
species <- read_csv("data/data_species_check_list.csv") %>% 
  rename(scientific_name = speciesname) %>% 
  distinct(scientific_name, .keep_all = TRUE) %>% 
  filter((annex_II_priority != "N" | annex_II != "N" | annex_IV != "N")) %>% 
  right_join(data, by = c("scientific_name")) %>% 
  drop_na(annex_II) %>% 
  count(scientific_name, common_name, annex_II_priority, annex_II, annex_IV) %>% 
  arrange(scientific_name)

kbl(species, booktabs = TRUE) %>% 
    kable_classic("hover", full_width = F, html_font = "Helvetica Neue")
```
